Modal Abstractions for Virtualizing Memory Addresses

### 2023-04-10 examples.tex
+

;; Ensure there's a place to put the L1 mapping
mov %rdi, %r13

@@

@@

@@

@@

@@

\caption{Distilled code for mapping a new page}

@@

\begin{figure}
\begin{lstlisting}
;; Assume the save-space is in rdi,
;; load-space in rsi
;; No need to save/restore caller-save regs
;; Save yielding context
mov 0[%rdi], %rbx
mov 8[%rdi], %rsp
mov 16[%rdi], %rbp
mov 24[%rdi], %r12
mov 32[%rdi], %r13
mov 40[%rdi], %r14
mov 48[%rdi], %r15
mov 56[%rdi], %cr3
;; Restore target context
mov %rbx, 0[%rsi] 
;; This switches to a new stack
;; This stack may not be mapped in the
;; current address space!
mov %rsp, 8[%rsi] 
mov %rbp, 16[%rsi]
mov %r12, 24[%rsi]
mov %r13, 32[%rsi]
mov %r14, 40[%rsi]
mov %r15, 48[%rsi]
;; Switch to the new address space
mov %cr3, 56[%rsi]
\end{lstlisting}
\caption{Basic task switch code that switches address spaces.}
\label{fig:swtch}
\end{figure}

>>>>

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred)  \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast \ulcorner \texttt{aligned va} \land  \theta \; !!\; \texttt{maddr} = \texttt{None}\urcorner}_{\rtv}$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \_ \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast }_{\rtv}$
$\specline{\exists (\entryf ,\;\entrytr,\; \entrytw,\; \entryo: \texttt{word 64}) \ldotp \ulcorner (\texttt{addr\_L1 }\vaddr) = \entryo\urcorner \ast }_{\rtv}$
$\specline{\nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}_{\rtv}$
$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast }_{\rtv}$
$\specline{ \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}_{\rtv}$
$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast}_{\rtv}$
$\specline{ \nfpointsto{\mask\vaddr\ft\rtv}{\mask\vaddr\tw\rtv}\entryf\qone\naddr \ast}_{\rtv}$ 
$\specline{  \nfpointsto{\mask\vaddr\ft\entryf}{\mask\vaddr\tw\entryf}\entrytr\qtwo\naddr \ast \nfpointsto{\mask\vaddr\ft\entrytr}{\mask\vaddr\tw\entrytr}\entrytw\qthree\naddr \ast}_{\rtv}$
$\specline{\texttt{pte\_addr} \mapsto_{\texttt{vpte},\rtv} \entryo \;(\texttt{wzero 64}) \ast \texttt{rax}\mapsto_{r} \texttt{pte\_addr}  }_{\rtv}$
$\specline{\exists \texttt{ fpaddr} \ldotp \ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$

@@

$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ \ast}_{\rtv}$
$\specline{\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\vaddr,\rtv,\entryf,\entrytr,\entrytw,\entryo) \ast \ulcorner \theta \;!!\; \texttt{maddr} = \texttt{None}\urcorner}_{\rtv}$
$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{fpaddr} \mapsto_{a} (\texttt{wzero 64})}_{\rtv}$
$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta<[\vaddr:=\texttt{fpaddr}]>),\gammaPred,\rtv) \ast  \texttt{r14}\mapsto_{r} \texttt{pte\_addr} \ast \texttt{rdi}\mapsto_{r} \vaddr \ast \texttt{rax}\mapsto_{r} \_ }_{\rtv}$
$\specline{\ulcorner \texttt{aligned fpaddr} \urcorner \ast \texttt{rax}\mapsto_{r} \texttt{fpaddr} \ast \texttt{aligned va} \ast\sumwalkabs\vaddr\qfrac\fpaddr}_{\rtv}$

@@

  \caption{Specification and proof of distilled code for mapping a new page with $\textsf{wpd\_def }_{\rtv}$}
  \todo[inline,color=yellow]{Colin, could you speak about the fact that we have wpd in iProp so that's the reason to have rtval whereever table traversal is needed}

@@

;; Assume the save-space is in rdi,
;; load-space in rsi
;; No need to save/restore caller-save regs
;; Save yielding context
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
$\specline{ \texttt{rsi}\mapsto_{r} \rtv' \ast \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+8} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+16} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+24} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+32} \mapsto_{v,\rtv} \_}_{\rtv}$
$\specline{\texttt{rdi+40} \mapsto_{v,\rtv} \_ \ast \texttt{rdi+48} \mapsto_{v,\rtv} \_\ast \texttt{rdi+56} \mapsto_{v,\rtv} \_}_{\rtv}$
mov 0[%rdi], %rbx
mov 8[%rdi], %rsp
mov 16[%rdi], %rbp
mov 24[%rdi], %r12
mov 32[%rdi], %r13
mov 40[%rdi], %r14
mov 48[%rdi], %r15
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast}_{\rtv}$
$\specline{ \texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \_}_{\rtv}$
mov 56[%rdi], %cr3
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \texttt{rbxv} \ast  \texttt{rsp}\mapsto_{r} \texttt{rspv} \ast \texttt{rbp}\mapsto_{r} \texttt{rbpv} }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{r} \texttt{r12v} \ast \texttt{r13}\mapsto_{r} \texttt{r13v} \ast \texttt{r14}\mapsto_{r} \texttt{r14v} \ast \texttt{r15}\mapsto_{r} \texttt{r15v}}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast }_{\rtv}$
$\specline{\texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}_{\rtv}$    
;; Restore target context
mov %rbx, 0[%rsi] 
;; This switches to a new stack
;; This stack may not be mapped in the
;; current address space!
mov %rsp, 8[%rsi] 
mov %rbp, 16[%rsi]
mov %r12, 24[%rsi]
mov %r13, 32[%rsi]
mov %r14, 40[%rsi]
mov %r15, 48[%rsi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred,\rtv) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{r} \rtv'+8 \ast \texttt{rbp}\mapsto_{r} \rtv'+16 }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{r} \rtv'+24 \ast \texttt{r13}\mapsto_{r} \rtv'+32 \ast \texttt{r14}\mapsto_{r} \rtv'+40 \ast \texttt{r15}\mapsto_{r} \rtv'+48}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast }_{\rtv}$
$\specline{\texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}_{\rtv}$    
;; Switch to the new address space
mov %cr3, 56[%rsi]
$\specline{\textsf{P} \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred) \ast [\rtv'+56](\mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother})}_{\rtv}$
$\specline{ \texttt{rdi}\mapsto_{r} \_ \ast \texttt{rbx}\mapsto_{r} \rtv'+0 \ast  \texttt{rsp}\mapsto_{r} \rtv'+8 \ast \texttt{rbp}\mapsto_{r} \rtv'+16 }_{\rtv}$
$\specline{\texttt{r12}\mapsto_{r} \rtv'+24 \ast \texttt{r13}\mapsto_{r} \rtv'+32 \ast \texttt{r14}\mapsto_{r} \rtv'+40 \ast \texttt{r15}\mapsto_{r} \rtv'+48}_{\rtv}$
$\specline{\texttt{rdi+0} \mapsto_{v,\rtv} \texttt{rbxv} \ast \texttt{rdi+8} \mapsto_{v,\rtv} \texttt{rspv} \ast \texttt{rdi+16} \mapsto_{v,\rtv} \texttt{rbpv} \ast  \texttt{rdi+24} \mapsto_{v,\rtv} \texttt{r12v} \ast }_{\rtv}$
$\specline{\texttt{rdi+32} \mapsto_{v,\rtv} \texttt{r13v} \ast \texttt{rdi+40} \mapsto_{v,\rtv} \texttt{r14v} \ast \texttt{rdi+48} \mapsto_{v,\rtv} \texttt{r15v}\ast \texttt{rdi+56} \mapsto_{v,\rtv} \rtv}_{\rtv}$
$\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\;\; \sqsubseteq $
$\specline{ [\rtv](\texttt{Pcurrent}) \ast \mathcal{I}\texttt{ASpace}(\theta,\gammaPred') \ast \texttt{Pother}}_{\rtv'+56}$
\caption{Basic task switch code that switches address spaces.}
\label{fig:swtch}
